### 2020-02-14 response/response-demography-2.tex

The author(s) have considered and addressed reviewers’ comments and have
revised the paper thoughtfully. The changes to the presentation of the
results and added explanations are especially well done. There is much
to be liked in the revised version. While the revised paper is stronger
than the earlier version, there are still some issues that require
further attention.

\begin{enumerate}

\item The paper still lacks a stronger narrative on the motivation and
framework for the empirical work. Four reasons are mentioned as
motivations; but, with the exception of the first, all other reasons,
such as the impact of spacing on child and mother’s health, are not
really motivating the empirical work of this paper (they are rather
evidence for importance of spacing).

I have entirely rewritten the paper to focus on how and why examining
birth spacing is relevant, as described above. As part of that rewrite,
I now estimate the relationship between spacing and infant mortality,
mainly to understand whether the increased use of sex selection may
increase mortality risk. Although I still find the effects on the
mothers' health a fascinating question, I have decided not to discuss
this effect because of lack of space.

@@

I have entirely rewritten the paper to focus on how and why examining
birth spacing is relevant, as described above. As part of that rewrite,
I now estimate the relationship between spacing and infant mortality,
mainly to understand whether the increased use of sex selection may
increase mortality risk. Although I still find the effects on the
mothers' health a fascinating question, I have decided not to discuss
this effect because of lack of space.

\item The conceptual framework is not fully developed. The paper suggests
possible explanations at various places but they don’t add up to a
coherent narrative. For instance, it is not clear whether maternal
education acts through access for sex selective technology or preference
for sons. The core idea of the paper that sex selection changes the
relationship between spacing and son preference needs to be developed
further. There is significant amount of good empirical work, but without
proper development of framework it is difficult to evaluate the
contribution and understand the findings, and also to see its
contribution beyond the Indian case.

As an illustration, the paper looks at changes over time during which
the access the sex selection has become increasing difficult AND also
almost all groups (by education, religion…) have seen an decline in
fertility. How do we consider the impact of these on spacing, sex
selection and sex ratios? How do we put the findings in the context of
changing educational attainments of women? The paper doesn’t provide us
a framework to think about the broader changes during the time period
under consideration.

The second of the two new background sections draw together the
different theories on the relationship between maternal education, labor
force participation, birth spacing, and use of sex selection, and use
those to guide the hypothesized changes.

To address the broader changes in India, the first of the two new
background sections show how female education has increased, while
female labor force participation has decreased. I argue that the low
female labor force participation is an indication of
``Sanskritization''. The pool of educated women has expanded, but the
behavior of the group does not change much.

I also show that although fertility fell, it is higher than generally
acknowledged, precisely because longer birth intervals biases downward
our standard fertility measures.

@@

I also show that although fertility fell, it is higher than generally
acknowledged, precisely because longer birth intervals biases downward
our standard fertility measures.

\item I’m not fully convinced of the main conclusions of a “substantial
increase in the use of sex selection” or of an “an almost complete
reversal of the traditional spacing patterns” for urban educated women.
They don’t seem to be supported by the findings. For urban women this
seems to apply only for third spell. The substantial increase in sex
selection is also not evident from the findings for all groups or for
all spells. The paper does acknowledge the complexity, but at places
(including in the abstract) seems to stretch the interpretation beyond
what is shown in the findings. The conclusions are presented in way that
implies that they apply to different spells or groups. But the findings
are more limited and differ clearly by spells. This should be presented
accurately in the text.

That was, indeed, an overstatement of the results. I now discuss the
three possible outcomes---no sex selection, sex selection with no change
in relative order by sex composition, and sex selection with a change in
the order---and provide examples of each.

@@

That was, indeed, an overstatement of the results. I now discuss the
three possible outcomes---no sex selection, sex selection with no change
in relative order by sex composition, and sex selection with a change in
the order---and provide examples of each.

\item While I really liked the methodology and visual presentation of the
findings, in my reading, I don’t see the paper’s contribution to
methodology. Though the paper makes a claim for methodological novely, I
see it more as an adoption of a widely used technique to spacing rather
than a proposing a new technique. Despite the changes, the paper still
reads as a piece demonstrating a methodology rather than signficant
contribution to the demogrpahic literature.

I have tried to move away from pitching the paper as a methodological
contribution. As part of this, I shortened the empirical model
discussion and removed any mention in the ``Conclusion'' section.
I hope that the two new analytical sections and the two background sections 
make the paper more of a definite contribution to the literature.

\item Minor: It might be a disciplinary style, I would recommend cutting down
on the footnotes.

I have removed or incorporated into the text a large number of footnotes.

\item I would also not go into the effect of child spacing
in countries that are very dissimilar to the context of India, unless
there is a wider point being made.

I have removed this discussion.

\item Some points are not clear: for
instance, “On the other hand, increased reliability of access and
effectiveness of contraceptives can lead to shorter intervals between
births if women used to have longer spacing to avoid having too many
children by accident (Keyfitz, 1971; Heckman and Willis, 1976).”

This discussion is no longer in the paper because of the new conceptual
framework. If other parts are unclear, please let me know.
